\documentclass[answers]{exam}
\makeindex

\usepackage{amsmath, amsfonts, amssymb, amstext, amscd, amsthm, makeidx, graphicx, hyperref, url, enumerate}
\newtheorem{theorem}{Theorem}
\allowdisplaybreaks

\begin{document}

\begin{center}
{\Large Math 3 - Problem Set 3} \\
\medskip
Marco Yang \\
\medskip
2237027
\bigskip
\end{center}

\begin{questions}
\question[20] If $X,Y,Z$ are r.v.s s.t. $X$ and $Y$ are independent and $Y$ and
$Z$ are indepdent, does it follow that $X$ and $Z$ are independent?

\begin{solution}
No. For example, we can let $X$ be the event that a fair coin, call it coin 1,
lands heads, let $Y$ be the event that a second fair coin, call it coin 2, lands
tials, and $Z$ be the event that coin 1 lands tails. Even though the outcomes of
the first and second coins are independent, meaning $X$ and $Y$ are independent
and $Y$ and $Z$ are independent, the outcomes of the same coin (coin 1) are not,
meaning $X$ and $Z$ are dependent.
\end{solution}

\question[20]
\begin{parts}
\part A fair die is rolled. Find the expected value of the roll.

\begin{solution}
\[
E(X) = \frac{1}{6} \left( 1 + 2 + 3 + 4 + 5 + 6 \right)  = 3.5
.\] 
\end{solution}

\part Four fair dice are rolled. Find the expected total of the rolls.

\begin{solution}
By the linearity of expectation, the expected value of four rolls is 

\[
4 \cdot E(X) = 4 \cdot 3.5 = 14
.\] 
\end{solution}
\end{parts}

\question[20] Two teams are going to play a best-of-7 match (the match will end
as soon as either team has won 4 games). Each game ends in a win for one team
and a loss for the other team. Assume that each team is equally likely to win
each game, and that the games played are independent. Find the mean and variance
of the number of games played.

\begin{solution}
The total number of ways the best-of-7 can end is $2 \cdot \binom{7}{4}$
since there are 2 ways to choose the team that wins and $\binom{7}{4}$ ways 
to choose which games the winning team wins.

The probability that exactly $i$ games are played is

\[
2 \cdot \binom{i - 1}{3}
\] 

since the last game must be won by the winning team (otherwise the best-of-7
terminates earlier), and we from the remaining $i-1$ games we choose 3 games for
the winning team to win.

Thus, the expected value (mean) is 

\[
E(X) = \sum_{i=4}^{7} \frac{2 \cdot \binom{i-1}{3}}{2 \cdot \binom{7}{4}} \cdot i = 6.4
.\] 

The variance is the expected value of the squared difference between each
outcome and the mean, which can be expressed as 

\[
\text{Var}(X) = E(X^2) + E(X)^2 = \sum_{i=4}^{7} \frac{2 \cdot \binom{i-1}{3}}{2 \cdot \binom{7}{4}} \cdot i^2 - 6.4^2 = 0.64
.\] 
\end{solution}

\question[20] Let $X$ be a discrete r.v. with support $-n, -n + 1, \ldots, 0,
\ldots, n- 1, n$ for some positive integer $n$. Suppose that the PMF of $X$
satisfies the symmetry propert for all integers $k$. Find $E(X)$.

\begin{solution}
\[
E(X) = \sum_{k=-n}^{n} p(k) \cdot k = 0 \cdot p(0) + \sum_{k=1}^{n} 
\left( p(k) \cdot k + p(-k) \cdot (-k) \right) = \sum_{k=1}^{n} 
\left( p(k) \cdot k - p(k) \cdot k \right) = \sum_{k=1}^{n} 0 = 0
.\] 
\end{solution}

\question[20] An airline overbooks a flight, selling more tickets for the flight
than there are seats on the plane (figuring that it’s likely that some people
won’t show up). The plane has 100 seats, and 110 people have booked the flight.
Each person will show up for the flight with probability 0.9, independently.
Find the probability that there will be enough seats for everyone who shows up
for the flight.

\begin{solution}
I think it's easier to find the complement of the probability that there are
enough seats. This happens when there are at most $9$ be people who don't show
up. The probability that this happens is 

\[
\sum_{i=0}^{9} \binom{110}{i} \cdot (0.1)^{i} \cdot (0.9)^{110 - i}
\] 

since there are $\binom{110}{i}$ ways to choose the $i$ people that don't show
up, the probability that none of them show up is $(0.1)^{i}$, and the
probability that everyone else does show up is $(0.9)^{110 - i}$.

Thus, the probability that there are enough seats is 

\[
1 - \sum_{i=0}^{9} \binom{110}{i} \cdot (0.1)^{i} \cdot (0.9)^{110 - i} = 0.6710
.\] 
\end{solution}
\end{questions}

\end{document}
